---
title: "Data Management and Web Scraping Workshop"
author: "YOUR NAME HERE"
date: "July 19, 2016"
output: html_document
---

# Case Study 1: NHANES
- In order to read in SAS files, we have to use a function from `Hmisc` called `sasxport.get`. 

```{r}
library(Hmisc)
demographics <- sasxport.get("http://wwwn.cdc.gov/Nchs/Nhanes/2013-2014/DEMO_H.XPT")
# copy the syntax used above to import the other dataset: Taste and Smell, file called "CSX_H.XPT"

```

## Clean Up, Clean Up, Everybody Everywhere...
1. Decide which variables to use
2. Look for missing data: why are they missing and is this a problem?
3. Look for out-of-range data: why would a living person have a pulse of 0 bpm?
4. Make sure numbers are numbers and factors are factors

```{r}
library(plyr)
library(dplyr)
# this is step 1; remember to rename the variables you select to something you'll remember
demo <- demographics %>% 
  select()
```

```{r}
# look for missing data, out-of-range data and data using summary
summary()
```

Is anything missing? Why? Is it a problem?
Is anything out-of-range? Is it an error or is the data real, yet confusing?
Are all of your factors reading as factors or are some of them numbers?

## More dplyr!

```{r}
# try using different dplyr functions here
# mutate

# filter

# group_by

# summarize

```

## Joining datasets
https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

```{r}

```

# Case Study 2: NBA
